---
title: ""
---

The general structure of a function in R is:

```{r}
function_name <- function(arg1, arg2 = default_value) {
  # body: operations using the arguments
  return(output)
}
```

A function definition has three main components:

* *Name:* the label you assign to your new function.
* *Arguments:* the inputs your function takes; they can have default values.
* *Body:* the code that executes when the function is called.

The `return()` statement specifies what the function should output, but it is optional. If omitted, R automatically returns the value of the last evaluated expression.

A simple example:

```{r}
square <- function(x) {
  x^2
}

square(4)
```

This function takes one argument (x), squares it, and returns the result. You can store the output or use it directly in a calculation:

```{r}
square(3) + 2
```

Functions are objects like any other in R—you can assign, print, or inspect them:

```{r}
square
```

The primary reason to write your own functions is to avoid repetition. Whenever you find yourself copying and pasting the same few lines of code, that’s a sign you should probably write a function.

Functions help you:

* Make code reusable: one definition can be called multiple times.
* Make code readable: meaningful names describe what the code does.
* Encapsulate complexity: you can hide the details of a process behind a simple interface.

For instance, suppose you frequently rescale numeric variables to a 0–1 range. Instead of repeating the logic each time, you can define a function once and call it whenever needed:

```{r}
rescale <- function(x, na.rm = TRUE) {
  rng <- range(x, na.rm = na.rm)
  (x - rng[1]) / (rng[2] - rng[1])
}
rescale(c(1, 5, 10))
```

This function now performs a well-defined transformation in a single, descriptive call.

::: practice

## Practice

Write a function that takes a temperature in Fahrenheit as input and returns the equivalent temperature in Celsius. The formula to convert Fahrenheit to Celsius is: C = (F - 32) * 5/9

:::: {.panel-tabset}
## {{< var tab_title.editor >}}

```{webr-r}

```


## {{< var tab_title.solution >}}

```{r}
fahrenheit_to_celsius <- function(f) {
  (f - 32) * 5 / 9
}
```
::::

:::
